---
title: "FIFA 2017"
author: "Nishanth Gandhidoss, Raghavendran Shankar"
date: "9/13/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```


```{r installing necessary packages, include=FALSE}
# installing the packages
installNewPackage <- function(packageName) {
        if(packageName  %in% rownames(installed.packages()) == FALSE)
        {
                install.packages(packageName, repos = "http://cran.us.r-project.org", dependencies=TRUE)
        }
}

installNewPackage("Amelia")
installNewPackage("wordcloud")
installNewPackage("caret")
installNewPackage("corrplot")
installNewPackage("moments")
installNewPackage("e1071")

library(Amelia)
library(wordcloud)
library(caret)
library(corrplot)
library(moments)
library(e1071)
```


## FIFA 2017

Something about the data - read the kaggle page

## Load the data

```{r loadthedata}
fifa_2017 <- read.csv(file = "data/FullData.csv", na.strings=c("","NA"))
head(fifa_2017)
```

## Dimensions of the Data

```{r dimension}
dim(fifa_2017)
```

Data has 17588 samples with 53 variables in it.

## Variables Word cloud

Let us visualize the word cloud for the whole variables in our dataset.

```{r word cloud}
# Setting the seed
set.seed(1)

# Getting the column names
col_name <- colnames(fifa_2017)

# Generate random numbers for frequency for the word cloud
col_name_values <- runif(length(col_name), min = 0, max = 1)

# Create the Word cloud for all variables
word_cloud <- data.frame(col_name, col_name_values)
wordcloud(word_cloud$col_name, freq = word_cloud$col_name_values, random.order = TRUE, random.color = FALSE, scale = c(2, 0.05), col = brewer.pal(9, "Oranges"))
```

As we have seen the whole variable word cloud let us see the continuous and categorical variable's word cloud.

```{r wordcloudi}
# Setting the seed
set.seed(1)

# Filtering the dataset
subset_colclasses <- function(DF, colclasses="numeric") {
  DF[,sapply(DF, function(vec, test) class(vec) %in% test, test=colclasses)]
}
fifa_2017_factor <- subset_colclasses(fifa_2017, c("factor"))
fifa_2017_integer <- subset_colclasses(fifa_2017, c("integer", "num"))

# Getting the column names
col_name_factor <- colnames(fifa_2017_factor)
col_name_integer <- colnames(fifa_2017_integer)

# Generate random numbers for frequency for the word cloud
col_name_factor_values <- runif(length(col_name_factor), min = 0, max = 1)
col_name_integer_values <- runif(length(col_name_integer), min = 0, max = 1)

# Create the Word cloud for all variables
word_cloud_factor <- data.frame(col_name_factor, col_name_factor_values)
wordcloud(word_cloud_factor$col_name_factor, freq = word_cloud_factor$col_name_factor_values, random.order = TRUE, random.color = FALSE, 
          scale = c(3.4, 0.5), col = brewer.pal(9, "Oranges"))
word_cloud_integer <- data.frame(col_name_integer, col_name_integer_values)
wordcloud(word_cloud_integer$col_name_integer, freq = word_cloud_integer$col_name_integer_values, random.order = TRUE, random.color = FALSE,
          scale = c(2, 0.5), col = brewer.pal(9, "Oranges"))
```


## Goal of the study

Now we are going to describe the goal of our study.

- Predicting the player rating
- Player field position

## Missing values in the data

Regarding the missing values in the predictors, the below image shows the missing values with the white places in the plot. 

```{r missingmap}
# Missmap
missmap(fifa_2017, col = c("white", "brown3"), main = "Missing Map | Before Processing")
```

The above map shows the missing values in white.

```{r missingcount1}
missing_count <- sapply(fifa_2017, function(x) sum(is.na(x)))
print(missing_count)
```

Thus we have the column names and number of missing values in each of the columns. To understand things more clear, lets filter out only those column having missing values leaving others.

```{r missingcount2}
# Get the missing variables alone
missing_count[missing_count > 0]
```


## Imputation of missing values

The Variables National_Position, National_Kit, Club_Kit, Club_joining, Contract_Expiry, Club_Position have atleast one missing value in the data. 
As the number of outliers is large in Club_fit and Contract_Expiry, the missing values are imputed by median value, The variables National_Position and National_kit are removed as they are not wanted.
The Club_position and Club_Joining have missing categorical values which are imputed by mode.

```{r Imputation of missing values}
# Remove unwanted predictors
fifa_2017$National_Position <- NULL
fifa_2017$National_Kit <- NULL

# Impute values by median
fifa_2017$Club_Kit = ifelse(is.na(fifa_2017$Club_Kit),ave(fifa_2017$Club_Kit, FUN = function(x) median(x, na.rm = TRUE)),fifa_2017$Club_Kit)
fifa_2017$Contract_Expiry = ifelse(is.na(fifa_2017$Contract_Expiry),ave(fifa_2017$Contract_Expiry, FUN = function(x) median(x, na.rm = TRUE)),fifa_2017$Contract_Expiry)

# Impute categorical values by mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(table(match(x, ux)))]
}

fifa_2017$Club_Position = ifelse(is.na(fifa_2017$Club_Position), ave(fifa_2017$Club_Position, FUN = function(x) Mode(x)), fifa_2017$Club_Position)
fifa_2017$Club_Joining = ifelse(is.na(fifa_2017$Club_Joining), ave(fifa_2017$Club_Joining, FUN = function(x) Mode(x)),fifa_2017$Club_Joining)
```

## Conversion of Height and Weight to numerical values

The height and weight variable are converted to numerical values

```{r Conversion to numerical values}
fifa_2017$Height <- as.numeric(gsub(fifa_2017$Height,pattern = " cm",replacement = ""))
fifa_2017$Weight <- as.numeric(gsub(fifa_2017$Weight,pattern = " kg",replacement = ""))
```

## Missing values verification of data

The data after imputation is verified for missing values. It is found that there is no missing value in the predictors.

```{r missmap}
missmap(fifa_2017, col = c("white", "darkolivegreen3"),main = "Missing Map | After Processing")
```

## Skewness

The Skewness for numerical predictors are found. The highly skewed variable (skewness > 1) is plotted in histogram

```{r Skewness greater than 1}
# Intialize variable
skew = list()

# Selecting only numerical values
fifa_num = sapply(X = fifa_2017,is.numeric)

# Calculating Skewness for the variables
skew = round(abs(apply(fifa_2017[,fifa_num], 2, skewness)), 3)
high_skew <- list()
high_skew_value <- list()
par(mfrow=c(3,3))

# Select skewness values greater than 1 and plot histogram
for(i in 1:length(skew)){
    if(skew[i] > 1){
        high_skew <- c(high_skew, names(skew[i]))
        high_skew_value <- c(high_skew_value, as.numeric(skew[i]))
        hist(fifa_2017[,names(skew[i])],main = paste(names(skew[i]), "|", skew[i]), cex.main = 3, xlab = names(skew[i]))
    }
}
```

## Box Cox Transformation

The Box Cox transformation is used to normalize the highly skewed predictos. Here, the highly skewed predictors are Club_kit, Ball_control, Dribbling, Short_Pass, Gk_Positioning, Gk_Diving,Gk_Kicking,Gk_Handling and Gk_Reflexes

```{r Removing skewness by Box Cox}
# Preprocessing the data and applying Box Cox transformation for highly skewed variables
fifa_trans <- preProcess(x = fifa_2017[unlist(high_skew)],method = "BoxCox")
fifa_data <- predict(fifa_trans,newdata = fifa_2017[unlist(high_skew)])

# Calculating Skewness after Box Cox transformation and Plotting histogram for the transformed variables
skew_measures <- matrix()
par(mfrow = c(3,3))
for (i in 1:ncol(fifa_data)){
    skew_measures[i] <- round(skewness(fifa_data[,i]), 3)
    hist(fifa_data[,i], main = paste(colnames(fifa_data)[i], "|", skew_measures[i]), cex.main = 3, xlab = colnames(fifa_data)[i]  )
}

data.frame(cbind(column = colnames(fifa_data), skew_measures))
```


## Correlation between Continuous predictors 

THe relationship between the predictors can be understood by seeing the correlation between the predictors. We can calculate the correlation and plot it visualize those. By this way, we can remove the one of the two predictors which are highly correlated with each other. Let see the correaltion first.

```{r corrplot}
par(xpd = TRUE)

# Plot the correlation plot 
corrplot(cor(fifa_2017_integer), order="hclust", tl.cex = 1)

# Finding out which predictors to elliminate since they have too large correlations
highCorr = findCorrelation(cor(fifa_2017_integer), cutoff = 0.90)
fifa_2017_integer_rem <- fifa_2017_integer[,-highCorr]
removed_col <- names(fifa_2017_integer)[highCorr]

# Matrix has no values > cutoff=0.9
corrplot(cor(fifa_2017_integer_rem), order="hclust", tl.cex = 1)

# Print the list of variables removed
print("The variables that are removed after performing correlation check are")
print(removed_col)
```



