---
title: "FIFA 2017"
author: "Nishanth Gandhidoss, Raghavendran Shankar"
date: "9/13/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```


```{r installing necessary packages, include=FALSE}
# installing the packages
installNewPackage <- function(packageName) {
        if(packageName  %in% rownames(installed.packages()) == FALSE)
        {
                install.packages(packageName, repos = "http://cran.us.r-project.org", dependencies=TRUE)
        }
}

installNewPackage("Amelia")

library(Amelia)
```


## FIFA 2017

Something about the data - read the kaggle page

## Load the data

```{r loadthedata}
fifa_2017 <- read.csv(file = "data/FullData.csv", na.strings=c("","NA"))
head(fifa_2017)
```

## Dimensions of the Data

```{r dimension}
dim(fifa_2017)
```

Data has 17588 samples with 53 variables in it.

## Goal of the study

- Predicting the player rating
- Player field position

## Missing values in the data

Regarding the missing values in the predictors, the below image shows the missing values with the white places in the plot. 

```{r missingmap}
# Missmap
missmap(fifa_2017, col = c("white", "red"))
```

The above map shows the missing values in white.

```{r missingcount}
sapply(fifa_2017, function(x) sum(is.na(x)))
```

Missing values in the data.